An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2017-02-24 uai2017/paper.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ICML 2017 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}
\usepackage{times,graphicx,subfigure,natbib,algorithm,algorithmic,hyperref}
\usepackage{amsmath,amsthm,amssymb,enumitem,verbatim}
% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\newcommand{\simiid}{\overset{\textrm{i.i.d.}}{\sim}}
\newcommand{\mE}{\mathbb{E}}
\DeclareMathOperator*{\argmin}{arg\,min}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\graphicspath{{figures/}{../figures/}}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
\usepackage{icml2017}

% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Proceedings of the...''
%\usepackage[accepted]{icml2017}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Submission and Formatting Instructions for ICML 2017}

@@

\twocolumn[
\icmltitle{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}

% It is OKAY to include author information, even for blind submissions: the
% style file will automatically remove it for you unless you've provided the
% [accepted] option to the icml2017 package.

% list of affiliations. the first argument should be a (short) identifier you
% will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% you can specify symbols, otherwise they are numbered in order
% ideally, you should not use this facility. affiliations will be numbered
% in order of appearance and this is the preferred way.
% \icmlsetsymbol{equal}{*}
\begin{icmlauthorlist}
\icmlauthor{Daniel Seita}{berkeley}
\icmlauthor{Xinlei Pan}{berkeley}
\icmlauthor{Haoyu Chen}{berkeley}
\icmlauthor{John Canny}{berkeley,google}
\end{icmlauthorlist}
\icmlaffiliation{berkeley}{University of California, Berkeley, CA}
\icmlaffiliation{google}{Google Research, Mountain View, CA}
\icmlcorrespondingauthor{Daniel Seita}{seita@berkeley.edu}

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boring formatting information, machine learning, ICML}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
%\footnotetext{hi}

@@

\section{Introduction}\label{sec:introduction}

@@

\section{Preliminaries}\label{sec:related_work}

@@

\subsection{A Worst-Case Gaussian Example}\label{ssec:gaussian_example}

@@

\subsection{MCMC Posterior Inference}
There is a separate line of MCMC work drawing principles from statistical
physics. One can apply
Hamiltonian Monte Carlo (HMC)~\cite{mcmc_hamiltonian_2010} methods which
generate high acceptance \emph{and} distant proposals when run on full batches
of data. Recently Langevin Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} has
been applied to Bayesian estimation on minibatches of data. This simplified
dynamics uses local proposals and avoids M-H tests by using small proposal steps
whose acceptance approaches 1 in the limit. However, the constraint on proposal
step size is severe, and the state space exploration reduces to a random walk.
Full minibatch HMC for minibatches was described in~\cite{sghmc_2014} which
allows momentum-augmented proposals with larger step sizes. However, step sizes
are still limited by the need to run accurately without M-H tests.  By providing
an M-H test with similar cost to standard gradient steps, our work opens the
door to applying those methods with much more aggressive step sizes without loss
of accuracy.

\section{A New MH Acceptance Test}\label{sec:our_algorithm}

\subsection{Log-Likelihood Ratios}\label{ssec:log_likelihood_ratios}

@@

\subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
For our new MH test we use the Barker logistic~\cite{Barker65}
function: $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows
that it satisfies the condition in Lemma~\ref{lem:detailed_balance}.
It is slightly less efficient than the Metropolis test, since its
acceptance rate for vanishing likelihood difference is 0.5. However we
will see that its overall sample efficiency is much higher than the
earlier methods.

@@

\subsection{A Minibatch Acceptance Test}\label{ssec:deltas_minibatch}

@@

\section{Computing the Correction Distribution}\label{sec:correction}

@@

\section{Analysis}\label{sec:analysis}

@@

\subsection{Bounding the Error of $\Delta^*$ from Gaussian}\label{ssec:delta_star_distribution}

@@

% Daniel: Lame, I know, but it saves one line.
%\subsection{Error Bounds are Preserved After Adding Random Variables}\label{ssec:preserve_bounds}
\subsection{Adding Random Variables}\label{ssec:preserve_bounds}

@@

\subsection{Bounds on the Stationary Distribution}
Bounds on the error of an M-H test imply bounds on the stationary distribution
of the Markov chain under appropriate conditions. Such bounds were derived in
both~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14}. We include the
result from~\cite{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the
total variation distance between two distributions $P$ and $Q$.  Let
$\mathcal{T}_0$ denote the transition kernel of the exact Markov chain,
$\mathcal{S}_0$ denote the exact posterior distribution, and
$\mathcal{S}_{\epsilon}$ denote the stationary distribution of the approximate
transition kernel. 
\begin{lemma}
If $\mathcal{T}_0$ satisfies the contraction condition
$d_v(P\mathcal{T}_0,\mathcal{S}_0) < \eta d_v(P,\mathcal{S}_0)$ for some
constant $\eta\in [0,1)$ and all probability distributions $P$, then
\begin{equation}
      d_v(S_0, S_{\epsilon}) \leq\frac{\epsilon}{1-\eta}
\end{equation}
where $\epsilon$ is the bound on the error in the acceptance test. 
\end{lemma}

\section{Experiments}\label{sec:experiments}

\subsection{Mixture of Gaussians}\label{ssec:gaussians}

@@

\subsection{Logistic Regression}\label{ssec:logistic}

@@

\begin{table}[t]
\caption{Log. Regression Average Minibatch Size (10 Trials)}
\label{tab:logistic}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}
\begin{tabular}{l l l l}
\hline
\abovespace\belowspace
Method/Data & MNIST-13k & MNIST8M-100k  \\
\hline
\abovespace
Our Method & $129.8\pm 3.4$ & $2059.9\pm 135.5$ \\
AustereMH  & $1924.3\pm 52.4$ & \textbf{TODO} \\
\belowspace
MHSubLhd   & $10783.4\pm 78.9$ & \textbf{TODO} \\
\hline
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

@@

\subsection{Collaborative Filtering}\label{ssec:collab_filtering}

@@

\section{Conclusions}\label{sec:conclusion}

@@

\bibliography{paper}
\bibliographystyle{icml2017}

@@

\section{Proofs of Lemmas and Corollaries}\label{app:proofs}

\subsection{Proof of Lemma~\ref{lem:worst_case}}\label{app:worst_case_proof}

@@

These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative by
inspection of equation (\ref{eq:whyneg}) below.  This implies that we can find a
$u \in (0,1)$ so that $\psi(u,\theta,\theta') = \log u < 0$ equals
$\mE[\Lambda^*(\theta,\theta')]$.  Specifically, choose $u_0$ to satisfy $\log u_0
= \mE[\Lambda^*(\theta,\theta')]$.  Using $\mE[x_i^*] = 0.5$ and
Equation~(\ref{eq:lemma_ll_ratio}), we see that
\begin{equation}
  \log u_0 = N(\theta'-\theta)\frac{1}{b} \cdot \mE\left[\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2}\right]
\end{equation}
\begin{equation}\label{eq:whyneg}
    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
\end{equation}
Next, consider the minibatch acceptance test $\Lambda^*(\theta,\theta')
\not\approx \psi(u,\theta,\theta')$ used in ~\cite{cutting_mh_2014}
and~\cite{icml2014c1_bardenet14} , where $\not\approx$ means ``significantly
different from'' under the distribution over samples. This is
\begin{align}
\Lambda^*(\theta,\theta')  \not\approx \psi(u_0,\theta,\theta') 
&\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2} \not\approx \log u_0\\
&\iff \frac{1}{b}\sum_{i=1}^b x_i^*-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) \not\approx  0 \\
&\iff \frac{1}{b}\sum_{i=1}^b x_i^*-0.5 \not\approx 0. \label{eq:accept_test_zero}
\end{align}
Since the $x_i^*$ have mean 0.5, the resulting test with our chosen $u_0$ will
never correctly succeed and must use all $N$ data points.  Furthermore, if we
sample values of $u$ near enough to $u_0$, the terms in parenthesis will not be
sufficiently different from 0.5 to allow the test to succeed.

@@

\subsection{Proof of Lemma~\ref{lem:quant_clt}}\label{app:quant_clt}

@@

\subsection{Proof of Lemma~\ref{lem:cdf_bounds}}\label{app:proof_cdf_bounds}

@@

\subsection{Proof of Corollary~\ref{cor:bounds_preserved}}\label{app:bounds_preserved}

@@

\subsection{Improved Error Bounds Based on Skew Estimation}\label{app:better_error_bound}

@@

\section{Additional Experiment Details}\label{app:experiments}
% Daniel: all three experiments should be described in more detail here. And
% also the deconvolution step to get the correction distribution.

\subsection{Obtaining the Correction Distribution (Section~\ref{sec:correction})}\label{app:correction}

% Daniel: moved to appendix so we can expand it. Originally we had only two
% lines, and it looked pretty awkward. So let's add more. Note: this also
% depends on whether we use [-10,10] for the X_c range, or [-20,20], etc...
\begin{table}[t]
\caption{Errors ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm
log}$, with $N=4000$ (top row) and $N=2000$ (bottom row).}
\label{tab:xcorr}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}
%% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=0.8$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 2.6e-3  \\
10     & 4.0e-4  \\
1      & 6.7e-5  \\
0.1    & 1.4e-5  \\
\belowspace
0.01   & \textbf{5.0e-6}  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=0.9$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 3.3e-3  \\
10     & 6.4e-4  \\
1      & 1.6e-4  \\
0.1    & \textbf{1.3e-4}  \\
\belowspace
0.01   & 2.7e-4  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=1.0$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 4.4e-3  \\
10     & 1.3e-3  \\
1      & \textbf{1.1e-3}  \\
0.1    & 2.0e-3  \\
\belowspace
0.01   & 3.6e-3  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=2000$ & $\sigma=1.1$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 6.8e-3  \\
10     & \textbf{4.6e-3}  \\
1      & 7.5e-3  \\
0.1    & 1.3e-2  \\
\belowspace
0.01   & 2.4e-2  \\
\hline
\end{tabular}
%% ---- NEW ROW ----
\vskip 0.2in %% OH this works!
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=0.8$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 8.3e-4  \\
10     & 1.3e-4  \\
1      & 2.5e-5  \\
0.1    & \textbf{6.7e-6}  \\
\belowspace
0.01   & 7.4e-6  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=0.9$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 1.2e-3  \\
10     & 2.6e-4  \\
1      & \textbf{1.0e-4}  \\
0.1    & 2.0e-4  \\
\belowspace
0.01   & 3.9e-4  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=1.0$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & 1.9e-3  \\
10     & \textbf{8.9e-4}  \\
1      & 1.6e-3  \\
0.1    & 2.8e-3  \\
\belowspace
0.01   & 5.2e-3  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
\abovespace\belowspace
$N=4000$ & $\sigma=1.1$ \\
\hline
\abovespace\belowspace
$\lambda$ & $L_{\infty}$ error \\
\hline
\abovespace
100    & \textbf{4.3e-3}  \\
10     & 6.0e-3  \\
1      & 1.0e-2  \\
0.1    & 1.2e-2  \\
\belowspace
0.01   & 3.5e-2  \\
\hline
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

@@

\subsection{Gaussian Mixture Model Experiment (Section~\ref{ssec:gaussians})}\label{app:gaussians}

\begin{table}[t]
\caption{Gaussian Mixture Model Statistics}
\label{tab:poissons}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}
\begin{tabular}{l l l l}
\hline
\abovespace\belowspace
Metric & Ours & Korat.'14 & Barde.'14 \\
\hline
\abovespace
Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
\belowspace
Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
\hline
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

@@

\subsection{Logistic Regression Experiment (Section~\ref{ssec:logistic})}\label{app:logistic}

@@

\subsection{Collaborative Filtering Experiment (Section~\ref{ssec:collab_filtering})}\label{app:collab_filtering}

@@

% Daniel: some notes to keep in mind to standardize to ICML format.

>>>>

% Sample LaTeX file for creating a paper in the Morgan Kaufmannn two
% column, 8 1/2 by 11 inch proceedings format.

\documentclass[letterpaper]{article}
\usepackage{proceed2e}
\usepackage[margin=1in]{geometry}
\usepackage{times,graphicx,subfigure,natbib,algorithm,algorithmic,hyperref}
\usepackage{amsmath,amsthm,amssymb,enumitem,verbatim}
% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\newcommand{\simiid}{\overset{\textrm{i.i.d.}}{\sim}}
\newcommand{\mE}{\mathbb{E}}
\DeclareMathOperator*{\argmin}{arg\,min}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\graphicspath{{figures/}{../figures/}}

\title{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}

% Daniel Seita: upon submission, comment out all authors, un-comment the
% following two lines:
%\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
%          % UAI  reviewing is double-blind.

% The author names and affiliations should appear only in the accepted paper.
%% \author{ {\bf Daniel Seita\thanks{Footnote for author to give an
%% alternate address.}} \\
%% Computer Science Dept. \\
%% Cranberry University\\
%% Pittsburgh, PA 15213 \\
%% \And
%% {\bf Xinlei Pan}  \\
%% Affiliation          \\
%% Address \\
%% \And
%% {\bf Haoyu Chen}   \\
%% Affiliation \\
%% Address    \\
%% \And
%% {\bf John Canny}   \\
%% Affiliation \\
%% Address    \\
%% }
\author{
Daniel Seita$^1$, 
Xinlei Pan$^1$,
Haoyu Chen$^1$,
John Canny$^{1,2}$ \\
$^1$ University of California, Berkeley, CA \\
$^2$ Google Research, Mountain View, CA\\ 
\texttt{\{seita,xinleipan,haoyuchen,canny\}@berkeley.edu}
}

@@

\maketitle

@@

\section{INTRODUCTION}\label{sec:introduction}

@@

\section{PRELIMINARIES}\label{sec:related_work}

@@

\subsection{A WORST-CASE GAUSSIAN EXAMPLE}\label{ssec:gaussian_example}

@@

\subsection{MCMC POSTERIOR INFERENCE}
There is a separate line of MCMC work drawing principles from statistical
physics. One can apply
Hamiltonian Monte Carlo (HMC)~\cite{mcmc_hamiltonian_2010} methods which
generate high acceptance \emph{and} distant proposals when run on full batches
of data. Recently Langevin Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} has
been applied to Bayesian estimation on minibatches of data. This simplified
dynamics uses local proposals and avoids M-H tests by using small proposal steps
whose acceptance approaches 1 in the limit. However, the constraint on proposal
step size is severe, and the state space exploration reduces to a random walk.
Full minibatch HMC for minibatches was described in~\cite{sghmc_2014} which
allows momentum-augmented proposals with larger step sizes. However, step sizes
are still limited by the need to run accurately without M-H tests.  By providing
an M-H test with similar cost to standard gradient steps, our work opens the
door to applying those methods with much more aggressive step sizes without loss
of accuracy.

\section{A NEW MH ACCEPTANCE TEST}\label{sec:our_algorithm}

\subsection{LOG-LIKELIHOOD RATIOS}\label{ssec:log_likelihood_ratios}

@@

\subsection{BARKER (LOGISTIC) ACCEPTANCE FUNCTION}\label{ssec:barker_function}
For our new MH test we use the Barker logistic~\cite{Barker65}
function: $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows
that it satisfies the condition in Lemma~\ref{lem:detailed_balance}.
It is slightly less efficient than the Metropolis test, since its
acceptance rate for vanishing likelihood difference is 0.5. However we
will see that its overall sample efficiency is much higher than the
earlier methods.

@@

\subsection{A MINIBATCH ACCEPTANCE TEST}\label{ssec:deltas_minibatch}

@@

\section{THE CORRECTION DISTRIBUTION}\label{sec:correction}

@@

\section{ANALYSIS}\label{sec:analysis}

@@

\subsection{BOUNDING THE ERROR OF $\Delta^*$ FROM A GAUSSIAN}\label{ssec:delta_star_distribution}

@@

% Daniel: Lame, I know, but it saves one line.
%\subsection{Error Bounds are Preserved After Adding Random Variables}\label{ssec:preserve_bounds}
\subsection{ADDING RANDOM VARIABLES}\label{ssec:preserve_bounds}

@@

\subsection{BOUNDS ON THE STATIONARY DISTRIBUTION}
Bounds on the error of an M-H test imply bounds on the stationary distribution
of the Markov chain under appropriate conditions. Such bounds were derived in
both~\cite{cutting_mh_2014} and~\cite{icml2014c1_bardenet14}. We include the
result from~\cite{cutting_mh_2014} (Theorem 1) here: Let $d_v(P,Q)$ denote the
total variation distance between two distributions $P$ and $Q$.  Let
$\mathcal{T}_0$ denote the transition kernel of the exact Markov chain,
$\mathcal{S}_0$ denote the exact posterior distribution, and
$\mathcal{S}_{\epsilon}$ denote the stationary distribution of the approximate
transition kernel. 
\begin{lemma}
If $\mathcal{T}_0$ satisfies the contraction condition
$d_v(P\mathcal{T}_0,\mathcal{S}_0) < \eta d_v(P,\mathcal{S}_0)$ for some
constant $\eta\in [0,1)$ and all probability distributions $P$, then
\begin{equation}
      d_v(S_0, S_{\epsilon}) \leq\frac{\epsilon}{1-\eta}
\end{equation}
where $\epsilon$ is the bound on the error in the acceptance test. 
\end{lemma}

\section{EXPERIMENTS}\label{sec:experiments}

\subsection{MIXTURE OF GAUSSIANS}\label{ssec:gaussians}

@@

\subsection{LOGISTIC REGRESSION}\label{ssec:logistic}

@@

\begin{table}[t]
\caption{Logistic Regression Average Minibatch Size}
\label{tab:logistic}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Method/Data} & \textbf{MNIST-13k} & \textbf{MNIST-100k}  \\
\hline \\
Our Method & $129.8\pm 3.4$ & $2059.9\pm 135.5$ \\
AustereMH  & $1924.3\pm 52.4$ & \textbf{TODO} \\
MHSubLhd   & $10783.4\pm 78.9$ & \textbf{TODO} \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

@@

\subsection{SPARSE MATRIX FACTORIZATION}\label{ssec:collab_filtering}

@@

\section{CONCLUSIONS}\label{sec:conclusion}

@@

\bibliography{paper}
\bibliographystyle{plain}

@@

\section{PROOFS OF LEMMAS AND COROLLARIES}\label{app:proofs}

\subsection{PROOF OF LEMMA~\ref{lem:worst_case}}\label{app:worst_case_proof}

@@

These events together guarantee that $\Lambda^*(\theta,\theta')$ is negative by
inspection of Equation (\ref{eq:whyneg}) below.  This implies that we can find a
$u \in (0,1)$ so that $\psi(u,\theta,\theta') = \log u < 0$ equals
$\mE[\Lambda^*(\theta,\theta')]$.  Specifically, choose $u_0$ to satisfy $\log u_0
= \mE[\Lambda^*(\theta,\theta')]$.  Using $\mE[x_i^*] = 0.5$ and
Equation~(\ref{eq:lemma_ll_ratio}), we see that
\begin{equation}\label{eq:whyneg}
    \log u_0 = N(\theta'-\theta)\frac{1}{b} \cdot \mE\left[\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2}\right]
    = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
\end{equation}
Next, consider the minibatch acceptance test $\Lambda^*(\theta,\theta')
\not\approx \psi(u,\theta,\theta')$ used in ~\cite{cutting_mh_2014}
and~\cite{icml2014c1_bardenet14} , where $\not\approx$ means ``significantly
different from'' under the distribution over samples. This is
\begin{align}
\Lambda^*(\theta,\theta')  \not\approx \psi(u_0,\theta,\theta') 
&\iff N(\theta'-\theta) \cdot \frac{1}{b}\sum_{i=1}^b x_i^*-\theta-\frac{\theta'-\theta}{2} \not\approx \log u_0\\
&\iff \frac{1}{b}\sum_{i=1}^b x_i^*-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right) \not\approx  0 \\
&\iff \frac{1}{b}\sum_{i=1}^b x_i^*-0.5 \not\approx 0. \label{eq:accept_test_zero}
\end{align}
Since the $x_i^*$ have mean 0.5, the resulting test with our chosen $u_0$ will
never correctly succeed and must use all $N$ data points.  Furthermore, if we
sample values of $u$ near enough to $u_0$, the terms in parenthesis will not be
sufficiently different from 0.5 to allow the test to succeed.

@@

\subsection{PROOF OF LEMMA~\ref{lem:quant_clt}}\label{app:quant_clt}

@@

\subsection{PROOF OF LEMMA~\ref{lem:cdf_bounds}}\label{app:proof_cdf_bounds}

@@

\subsection{PROOF OF COROLLARY~\ref{cor:bounds_preserved}}\label{app:bounds_preserved}

@@

\subsection{IMPROVED ERROR BOUNDS BASED ON SKEW ESTIMATION}\label{app:better_error_bound}

@@

\section{ADDITIONAL EXPERIMENT DETAILS}\label{app:experiments}
% Daniel: all three experiments should be described in more detail here. And
% also the deconvolution step to get the correction distribution.

\subsection{OBTAINING THE CORRECTION DISTRIBUTION (SECTION~\ref{sec:correction})}\label{app:correction}

%% % Daniel: moved to appendix so we can expand it. Originally we had only two
%% % lines, and it looked pretty awkward. So let's add more. Note: this also
%% % depends on whether we use [-10,10] for the X_c range, or [-20,20], etc...
\begin{table}[t]
\caption{Errors ($L_\infty$) in $X_{\rm norm}+X_{\rm corr}$ versus $X_{\rm
log}$, with $N=4000$ (top row) and $N=2000$ (bottom row).}
\label{tab:xcorr}
\vskip 0.15in
\begin{center}
\begin{small}
%% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=2000$ & $\sigma=0.8$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 2.6e-3  \\
10     & 4.0e-4  \\
1      & 6.7e-5  \\
0.1    & 1.4e-5  \\
0.01   & \textbf{5.0e-6}  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=2000$ & $\sigma=0.9$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 3.3e-3  \\
10     & 6.4e-4  \\
1      & 1.6e-4  \\
0.1    & \textbf{1.3e-4}  \\
0.01   & 2.7e-4  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=2000$ & $\sigma=1.0$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 4.4e-3  \\
10     & 1.3e-3  \\
1      & \textbf{1.1e-3}  \\
0.1    & 2.0e-3  \\
0.01   & 3.6e-3  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=2000$ & $\sigma=1.1$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 6.8e-3  \\
10     & \textbf{4.6e-3}  \\
1      & 7.5e-3  \\
0.1    & 1.3e-2  \\
0.01   & 2.4e-2  \\
\hline
\end{tabular}
%% ---- NEW ROW ----
\vskip 0.2in %% OH this works!
\begin{tabular}{l l}
\hline
$N=4000$ & $\sigma=0.8$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 8.3e-4  \\
10     & 1.3e-4  \\
1      & 2.5e-5  \\
0.1    & \textbf{6.7e-6}  \\
0.01   & 7.4e-6  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=4000$ & $\sigma=0.9$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 1.2e-3  \\
10     & 2.6e-4  \\
1      & \textbf{1.0e-4}  \\
0.1    & 2.0e-4  \\
0.01   & 3.9e-4  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=4000$ & $\sigma=1.0$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & 1.9e-3  \\
10     & \textbf{8.9e-4}  \\
1      & 1.6e-3  \\
0.1    & 2.8e-3  \\
0.01   & 5.2e-3  \\
\hline
\end{tabular}
\quad %% ----------------------- START OF TABLE
\begin{tabular}{l l}
\hline
$N=4000$ & $\sigma=1.1$ \\
\hline
$\lambda$ & $L_{\infty}$ error \\
\hline
100    & \textbf{4.3e-3}  \\
10     & 6.0e-3  \\
1      & 1.0e-2  \\
0.1    & 1.2e-2  \\
0.01   & 3.5e-2  \\
\hline
\end{tabular}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

@@

\subsection{GAUSSIAN MIXTURE MODEL EXPERIMENT (SECTION~\ref{ssec:gaussians})}\label{app:gaussians}

\begin{table}[t]
\caption{Gaussian Mixture Model Statistics}
\label{tab:poissons}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Metric} & \textbf{Ours} & \textbf{AustereMH} & \textbf{MHSubLhd} \\
\hline \\
Equation~\ref{eq:log_prob_poissons} & -1430.0 & -1578.9 & -1232.7 \\
Chi-Squared & 3313.9 & 3647.7 & 2444.1 \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

@@

\subsection{LOGISTIC REGRESSION EXPERIMENT (SECTION~\ref{ssec:logistic})}\label{app:logistic}

@@

\subsection{SPARSE MATRIX FACTORIZATION EXPERIMENT (SECTION~\ref{ssec:collab_filtering})}\label{app:collab_filtering}

@@
